---
title: "Coursera repdata-016: Project2"
output: html_document
---

## Synopsis
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

(cited from coursera course project instructions)

## R sessionInfo
```{r}
sessionInfo()
```

## Data Processing
download and load the raw data
```{r ,cache=TRUE}
# download data
library(RCurl)
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", destfile = 'data.csv.bz2',method='curl')
# extract csv data
#library(R.utils)
#bunzip2("data.csv.bz2",destname = 'data.csv', overwrite=TRUE, remove=F)
# read in data
data<-read.table(bzfile('data.csv.bz2'),header=T,sep=',')
dim(data)
names(data)
```

process the data  

* population health  
use "data\$FATALITIES+data\$INJURIES" as measure of population health
```{r ,cache=TRUE}
data$health = data$FATALITIES+data$INJURIES
health=aggregate(data$health,by=list(data$EVTYPE),sum,na.rm=T)
names(health)=c('type','num')
# select top 10 types
health=health[order(health$num,decreasing = T),]
health=health[1:10,]
print(health)
```

* economic  
use PROPDMG and CROPDMG as measure of economic loss
```{r ,cache=TRUE}
# process the raw PROPDMGEXP and CROPDMGEXP 
unique(data$PROPDMGEXP)
unique(data$CROPDMGEXP)
library(plyr)
data$propexpnum <- as.numeric(as.character((revalue(data$PROPDMGEXP, c("B"=9, "M" = 6, "m" = 6,"K" = 3,"H" = 2, "h" = 2,"-"= NA, "+" =NA, "?" = NA)))))
data$croexpnum <- as.numeric(as.character(revalue(data$CROPDMGEXP, c("B"=9, "M" = 6, "m" = 6,"K" = 3,"k" = 3, "?" =NA))))
data$econ = data$PROPDMG*10^data$propexpnum+data$CROPDMG*10^data$croexpnum
# top 10 types
econ=aggregate(data$econ,by=list(data$EVTYPE),sum,na.rm=T)
names(econ)=c('type','num')
# select top 10 types
econ=econ[order(econ$num,decreasing = T),]
econ=econ[1:10,]
print(econ)
```

## Results
```{r}
par(mar=c(12,4,4,2))
barplot(health$num,names.arg = health$type,main='population health',las=2)
barplot(econ$num,names.arg = econ$type,main='economic',las=2)
```  

So,
 - population health: tornado
 - economic consequences: flood




