---
title: "project of regmods-010"
output:
  pdf_document: default
---

## Executive summary
The goal of this project is to find out:
- “Is an automatic or manual transmission better for MPG”
- "Quantify the MPG difference between automatic and manual transmissions"

Results:
On a significant level of 99%, manual is better than automatic in terms of mpg.
Regression shows Manual is 2.083710 mpg better than Automatic.

## Data Processing

```{r}
library(datasets)
data(mtcars)
#str(mtcars)
# am is factor variable
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <- c("Automatic", "Manual")
```


## Exploratory Data Analysis
```{r,eval=FALSE}
# look at the y values
hist(mtcars$mpg,20,probability = T, main='histogram of mpg and fitted dnesity')
lines(density(mtcars$mpg))
# relation ship between y and other x values
pairs(mtcars)
```

In particular, how mpg differs in am.
```{r,eval=FALSE}
boxplot(mpg~am, data = mtcars,
        main = "mpg vs am")
```

Do a hypothesis testing
```{r}
t.test(mtcars$mpg[mtcars$am=='Automatic'], mtcars$mpg[mtcars$am=='Manual'])
``` 
On a significant level of 99%, manual is better than automatic in terms of mpg.


## Regression Analysis
```{r}
# simple linear regression
fit0 = lm(mpg~am, mtcars)
#summary(fit0)
# multiple regression
fit1 <- lm(mpg~am + wt + hp, data = mtcars)
#summary(fit1)
par(mfrow = c(2,2))
plot(fit1)
# anova test
anova(fit0, fit1)
```
From fit1 model, we can see that Manual is 2.083710 mpg better than automatic.

## Appendix
```{r,echo=FALSE}
boxplot(mpg~am, data = mtcars,
        main = "mpg vs am")
```

```{r,echo=FALSE}
# look at the y values
hist(mtcars$mpg,20,probability = T, main='histogram of mpg and fitted dnesity')
lines(density(mtcars$mpg))
# relation ship between y and other x values
pairs(mtcars)
```

